The Driving Forces of API Evolution

### 2015-07-13 thesis/paper.tex
Camera Ready
1. Changes based on feedback from reviewers

2. Switched order of authors (john+william)

\author{
% 1st. author
\alignauthor
John Burchell and William Granli\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{john.a.burchell@gmail.com},  \email{william.granli@gmail.com}
% 2nd. author
%\alignauthor
%William Granli\\
%       \affaddr{Computer Science and Engineering}\\
%       \affaddr{University of Gothenburg}\\
%       \email{william.granli@gmail.com}
%\and
% 3rd. author
\alignauthor Imed Hammouda and Eric Knauss\\
\affaddr{Computer Science and Engineering}\\
\affaddr{Chalmers $\mid$ University of Gothenburg}\\
\email{\{imed.hammouda,eric.knauss\}@cse.gu.se}
% 4th 
%\alignauthor Eric Knauss\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
%       \affaddr{University of Gothenburg}\\
%       \email{imed.hammouda@gu.se}
}

@@

\begin{abstract}
%Motivation
Evolving an Application Programming Interface (API) is a precipitous activity, as modifications to them can significantly impact their users. The increasing use of APIs means that software development organisations must take an empirical and scientific approach to the way they manage the evolution of their APIs. 
%Problem
If no attempt at analysing or quantifying the evolution of an API is made, there will be a diminished understanding of the evolution, and possible improvements to the maintenance strategy will be difficult to identify. We believe that long-standing software evolution theories can provide additional insight to the field of APIs, and can be of great use to companies maintaining APIs. 
%Method
In this case study, we conduct a qualitative investigation to understand what drives the evolution of an industry company's existing API, by examining two versions of the API code. The changes were analysed based on two software evolution theories, and the extent to which we could reverse engineer the change decisions was determined by interviewing an architect of the API. 
%Results
The results of this analysis show that the largest driving force of the APIs evolution was the desire for new functionality. Our findings which show that changes happen sporadically, rather than continuously, appear to show that the law of Conservation of Organisational Stability was not a considerable factor for the evolution of the API. We also found that it is possible to reverse engineer change decisions and in doing so, identified that the feedback loop of an API is an important area of improvement. 
\end{abstract}

@@

%Purpose, general
Modifications to deployed APIs may negatively impact its users. It is therefore important to investigate why these changes occur. Understanding the motives behind the changes could help API architects to prevent potential future changes to their APIs. Before industry can safely include such information in their feedback loop, we must critically assess to what degree we can reverse engineer such motives. 
%Purpose, case company
Identifying the types of changes that occur when an API evolves will improve communication and discussions regarding maintenance and evolution tasks for the case company. This improved communication could provide a common language that developers, users and management can use to discuss potential evolution of the APIs. Understanding the evolution theories could further assist understanding of how change will occur and how to plan for it. Lastly, by keeping a history of changes to an API over time, trends can be identified which could help with future API design.

%Gap in Literature
Previous studies have investigated programming language APIs, libraries or frameworks focusing on causes and affects of modifications to APIs \cite{dig2005role, hou2011exploring, shi2011empirical}. However, to the best of our knowledge, no studies have been performed on embedded platform APIs from a software evolution perspective. Our study will fill this gap by analysing an embedded platform API, classifying the types of changes that occur and analysing them with the use of software evolution theories.

@@

              \textbf{Type}                              & \textbf{Explanation}    \\ \midrule
              Corrective    & Corrections of functionality                   \\ \hline
              Reductive     & Reductions of functionality                    \\ \hline
              Adaptive      & Inclusion of new Technology                    \\ \hline
              Performance   & Improvements to Performance                    \\ \hline
              Preventive    & Improvements to future maintainability         \\ \hline
              Groomative    & Improvements to maintainability                \\ \hline
              Updative      & Updates to documentation                       \\ \hline
              Reformative   & Changes to documentation                       \\ \hline
              Evaluative    & Inspection activities                          \\ \hline
              Consultive    & Consultations activities                       \\ \hline
              Training      & Training activities                            \\

@@

\begin{table}
       \centering
       \begin{tabular}[ht]{p{3in}}
       \toprule
              \begin{enumerate}
              \item \textbf{Continuing Change}  \newline
              E-Type systems must be continually adapted else they become progressively less satisfactory.                     
              \item \textbf{Increasing Complexity} \newline
              As an E-Type system evolves its complexity increases unless work is done to maintain or reduce it.                  
              \item \textbf{Self Regulation} \newline
              E-Type system evolution process is self regulating with distribution of product and process measures close to normal.                     
              \item \textbf{Conservation of Organisational Stability} \newline
              The average effective global activity rate in an evolving E-Type system is invariant over product lifetime.                     
              \item \textbf{Conservation of Familiarity} \newline
              As an E-type system evolves all associated with it, developers, sales personnel, users, for example, must maintain mastery of its content and behaviour to achieve satisfactory evolution. Excessive growth diminishes that mastery. Hence the average incremental growth remains invariant as the system evolves.                     
              \item \textbf{Continuing Growth} \newline
              The functional content of E-Type systems must be continually increased to maintain user satisfaction over their lifetime.                
              \item \textbf{Declining Quality} \newline
              The quality of E-Type systems will appear to be declining unless they are rigorously maintained and adapted to operational environment changes.                     
              \item \textbf{Feedback System} \newline
              E-Type evolution process constitute multi-level, multi-loop, multi-agent feedback systems and must be treated as such to achieve significant improvement over any reasonable base.

              \end{enumerate} \\
              \bottomrule
       \end{tabular}
       \caption{Lehman's Laws of Software Evolution \cite{lehman90sview}}
       \label{table:expl_laws}
\end{table}

@@

\subsection{Case Company Description} \label{case_company_description}
The case company of this study will from here on be referred to as \textit{company A}. Company A operates in the domain of video surveillance. The company's headquarters is located in Lund, Sweden, with offices in 49 countries worldwide. Company A is the global market leader in the markets of network cameras and video encoders. They develop embedded software for security cameras. The cameras are designed to be accessible through APIs that are developed and maintained by the company.

@@

\section{Methodology} \label{methodology}
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis. The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, robson2002real, runeson2009guidelines, yin2013case}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of drivers of API change and that input from the industry is vital to the success of the research.

@@

The study was conducted in two phases. The aim of the first phase was to generate hypotheses and to formulate clear research questions. This was achieved by using a data-driven approach to analysing the code and documentation. The second phase included analysis of the interviews, during which we aimed to confirm the hypotheses that were formulated in the first phase. Both phases were performed iteratively. This created the opportunity to improve the data analysis as the study progressed, as well as, allowing us to adapt to possible changes in direction, due to the hypothesis generating approach of the first phase. An overview of the methodology can be found in Table \ref{table:meth_steps}.

@@

%How the initial inspection was performed
The API code was inspected sequentially, starting with v1.4. Both versions of the API underwent the same type of inspection. This involved extracting method signatures, enums, structs, typedefs and macros, which were subsequently stored in spreadsheets. The extraction involved a lot of manual work, but Git's \cite{git} diff command was used to identify changes in files which exist in both versions. After a change was identified in the code, a description of the corresponding change in the documentation was added to the spreadsheet. The module and file in which the change was identified was then logged together with an ID and a description of the change. The data collection from code and documentation was performed independently by both researchers to reduce the risk of human errors affecting the results. If the data collected by each researcher differed, the cause of the discrepancy was investigated and resolved.

@@

%Codes
To analyse the data, each change identified in the spreadsheet was coded based on which module the change occurred in and two other characteristics: a) if the change added, removed or modified functionality, and b) if the change was a cosmetic change or not.
%Concepts
Grouping the codes by module allowed us to form general concepts of change. However, changes that affected multiple modules were classified on their own.
%Categories
These concepts were then grouped into categories based on common patterns between the concepts.
%Theory
These categories were then used to identify the most notable trends in the evolution of the API. These trends were later used as a basis for discussion during the interview.

%RQ1
The trends identified previously were then classified with help of the decision tree used for classifying the types of software evolution and software maintenance \cite{chapin2001types}. Each of the previously identified trends were given a main type of change and, if appropriate, a secondary type. Main types of change identify the primary motive for a trend, whereas the secondary types are additional motivations for a trend. For example, if an identified trend was the re-implementation of an old module that included new functionality and made use of a new interface, it would be classified as mainly enhancive and secondarily groomative. In addition to this, we analysed the trends based on Lehman's Laws, to determine if they applied to the changes made to the API. This information, as well as the results of the interview, were then used as the basis for answering \textbf{RQ1}.

@@

\smallskip \noindent
\textbf{Updative  }
%Define
Updative changes are defined to be changes to documentation of a system.
%Data and example
A single change was determined to be updative; updating the API documentation to include the new functionality of v2.0. The updates included a new API specification, deprecation list and the inclusion of the new features and libraries.
%Conclusion
Updative changes have very little impact on the functionality of the API yet they are still important changes for the usability of the API \cite{shi2011empirical}. Changes to the API cause the need for updative changes to its documentation, without them, the usability of the API could suffer. We therefore conclude that the evolution of this API was not driven by updative changes.

@@

\smallskip \noindent
\textbf{Reformative  }
%Define
Reformative changes are defined to be changes that update the documentation to the stakeholders needs.
%Data and Example
One reformative change was identified between v1.4 and v2.0; a redesign of the existing example programs. The aim of this change was to update the examples to illustrate the new functionality in the API, showing how the modules interact and at the same time, presenting a standard for how to use the API.
%Conclusion
We therefore conclude that reformative changes, while important for the usability of the API, did not drive the evolution of the API.

@@

\smallskip \noindent
\textbf{Conservation of Familiarity  } 
The documentation that is included with the API was rigorously updated with the introduction of v2.0. All functionality which was added or modified in v2.0 was updated accordingly in the documentation. The same pattern can be seen for the example code, which went through major changes as a result of the changes to the API code. Company A's intent to control the way that their API is used also suggests that the knowledge of how to use the API is of importance, since if the API is used in too many different ways, it will be more difficult for company A to have relevant documentation and example code. It was also expressed during the interview that further transparency in how the API should be used could be achieved. One example mentioned of how this could be achieved is by improving the expressiveness regarding which API functions are supported on which hardware types. These facts support that the law of Conservation of Familiarity played an important part in the evolution of the API.

@@

\smallskip \noindent
\textbf{Declining Quality  } 
The interviews established that a groomative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. Our analysis also shows that changes related to quality are bound to happen, as other types of changes often incorporate quality related aspects. We therefore settle that the law of Declining Quality was a driving force for the evolution of the API.

@@

\subsection{Reverse Engineering API Change Decisions}
After cross-examining our initial analysis with the interviewee's responses, we found that 85\% of our main type predictions matched the actual motive of the change and that 85\% of our secondary type predictions were correct. The kappa value \cite{cohen1968weighted} for the predictions was 0.812, which shows that there is a high level of agreement between the predictions and the change decisions. In Table \ref{table:reverse_engineering}, a detailed view of each prediction juxtaposed against the actual change decision can be found for both the main and secondary types.

@@

\subsection{Implications for Research} 
The implications of our findings should encourage research to further study APIs by relating them to existing software evolution theories. Considering that we were able to successfully reverse engineer the changes made to the API in this study, it opens up possibilities to base future research on, for example, the types of software evolution and software maintenance. Since the kappa analysis showed that there was a high level of agreement between our predictions and the actual motives to the change, it shows that the types of software evolution and software maintenance could even be used as the basis for a large-scale quantitative analysis to map the types of changes made to APIs.

@@

One way of finding the solution to this problem would be by including the application developers in the feedback loop. Since the application developers are on the receiving end of the changes, we believe that their needs have to be assessed before the problem can be solved. With our study we identified both that the law of Feedback System is of great importance to company A and that it is one of the key areas of API development that company A wishes to improve. One of the challenges posed by software evolution is that the activities carried out are often intangible and non-objective \cite{chapin2001types, lientz1980software}. Based on this, we propose that our study which has granulated the activities is used as a basis for communicating the current stage of the evolution of their API. We hope that this classification can be used in its current form, to help quantify the activities and to provide an overview of them. We also hope that the approach we used can be used to predict which types of activities should be emphasised in the future evolution of the API. With our study we have also demonstrated that the classification of changes is achievable, which, in turn, should encourage company A to adopt this method of classifying change.

\section{Conclusion} \label{conclusion}
With this study, we conclude that studying API evolution with help of existing theories, provides a new and structured approach to the field of API evolution. Our study shows that functionality is the largest driving force of change in the evolution of the studied API. It also suggests that changes to APIs occur infrequently and sporadically. This indicates that the law of Conservation of Organisational Stability did not play a significant role in the evolution of this API. However, given that this study was on performed on a single API, we cannot generalise this finding. Furthermore, we have identified the importance of having a structured feedback loop when maintaining an API. We have also provided company A with a foundation for facilitating such a change.

>>>>

\author{
% 1st. author
\alignauthor
William Granli and John Burchell\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{william.granli@gmail.com},  \email{john.a.burchell@gmail.com}
% 2nd. author
%\alignauthor
%William Granli\\
%       \affaddr{Computer Science and Engineering}\\
%       \affaddr{University of Gothenburg}\\
%       \email{william.granli@gmail.com}
%\and
% 3rd. author
\alignauthor Imed Hammouda and Eric Knauss\\
\affaddr{Computer Science and Engineering}\\
\affaddr{Chalmers $\mid$ University of Gothenburg}\\
\email{\{imed.hammouda,eric.knauss\}@cse.gu.se}
% 4th 
%\alignauthor Eric Knauss\titlenote{Co-author}\\
%       \affaddr{Computer Science and Engineering}\\
%       \affaddr{University of Gothenburg}\\
%       \email{imed.hammouda@gu.se}
}

@@

\begin{abstract}
%Motivation
Evolving an Application Programming Interface (API) is a delicate activity, as modifications to them can significantly impact their users. The increasing use of APIs means that software development organisations must take an empirical and scientific approach to the way they manage the evolution of their APIs. 
%Problem
If no attempt at analysing or quantifying the evolution of an API is made, there will be a diminished understanding of the evolution, and possible improvements to the maintenance strategy will be difficult to identify. We believe that long-standing software evolution theories can provide additional insight to the field of APIs, and can be of great use to companies maintaining APIs. 
%Method
In this case study, we conduct a qualitative investigation to understand what drives the evolution of an industry company's existing API, by examining two versions of the API interface. The changes were analysed based on two software evolution theories, and the extent to which we could reverse engineer the change decisions was determined by interviewing an architect of the API. 
%Results
The results of this analysis show that the largest driving force of the APIs evolution was the desire for new functionality. Our findings which show that changes happen sporadically, rather than continuously, appear to show that the law of Conservation of Organisational Stability was not a considerable factor for the evolution of the API. We also found that it is possible to reverse engineer change decisions and in doing so, identified that the feedback loop of an API is an important area of improvement. 
\end{abstract}

@@

%Purpose, general
Modifications to deployed APIs may negatively impact its users. It is therefore important to investigate why these changes occur. Understanding the motives behind the \newline changes could help API architects to prevent potential future changes to their APIs. Before industry can safely include such information in their feedback loop, we must critically assess to what degree we can reverse engineer such motives. 
%Purpose, case company
Identifying the types of changes that occur when an API evolves will improve communication and discussions regarding maintenance and evolution tasks for the case company. This improved communication could provide a common language that developers, users and management can use to discuss potential evolution of the APIs. Understanding the evolution theories could further assist understanding of how change will occur and how to plan for it. Lastly, by keeping a history of changes to an API over time, trends can be identified which could help with future API design.

%Gap in Literature
Previous studies have investigated programming language APIs, libraries or frameworks focusing on causes and effects of modifications to APIs \cite{dig2005role, hou2011exploring, shi2011empirical}. However, to the best of our knowledge, no studies have been performed on embedded platform APIs from a software evolution perspective. Our study will fill this gap by analysing an embedded platform API, classifying the types of changes that occur and analysing them with the use of software evolution theories.

@@

              \textbf{Type}                              & \textbf{Explanation}    \\ \midrule
              Enhancive     & Inclusion of new functionality                 \\ \hline
              Corrective    & Corrections of functionality                   \\ \hline
              Reductive     & Reductions of functionality                    \\ \hline
              Adaptive      & Inclusion of new technology                    \\ \hline
              Performance   & Improvements to performance                    \\ \hline
              Preventive    & Improvements to future maintainability         \\ \hline
              Groomative    & Improvements to maintainability                \\ \hline
              Updative      & Updates to documentation                       \\ \hline
              Reformative   & Changes to documentation                       \\ \hline
              Evaluative    & Inspection activities                          \\ \hline
              Consultive    & Consultations activities                       \\ \hline
              Training      & Training activities                            \\

@@

\begin{table}
       \centering
       \begin{tabular}[ht]{p{3in}}

              \begin{enumerate}
              \item \textbf{Continuing Change}  \newline
              E-Type systems must be continually adapted else they become progressively less satisfactory.                     
              \item \textbf{Increasing Complexity} \newline
              As an E-Type system evolves its complexity increases unless work is done to maintain or reduce it.                  
              \item \textbf{Self Regulation} \newline
              E-Type system evolution process is self regulating with distribution of product and process measures close to normal.                     
              \item \textbf{Conservation of Organisational Stability} \newline
              The average effective global activity rate in an evolving E-Type system is invariant over product lifetime.                     
              \item \textbf{Conservation of Familiarity} \newline
              As an E-type system evolves all associated with it, developers, sales personnel, users, for example, must maintain mastery of its content and behaviour to achieve satisfactory evolution. Excessive growth diminishes that mastery. Hence the average incremental growth remains invariant as the system evolves.                     
              \item \textbf{Continuing Growth} \newline
              The functional content of E-Type systems must be continually increased to maintain user satisfaction over their lifetime.                
              \item \textbf{Declining Quality} \newline
              The quality of E-Type systems will appear to be declining unless they are rigorously maintained and adapted to operational environment changes.                     
              \item \textbf{Feedback System} \newline
              E-Type evolution process constitute multi-level, multi-loop, multi-agent feedback systems and must be treated as such to achieve significant improvement over any reasonable base.

              \end{enumerate} \\

       \end{tabular}
       \caption{Lehman's Laws of Software Evolution \cite{lehman90sview}}
       \label{table:expl_laws}
\end{table}

@@

\subsection{Case Company Description} \label{case_company_description}
The case company of this study will from here on be referred to as \textit{company A}. Company A operates in the domain of video surveillance and has offices in 49 countries worldwide. Company A is the global market leader in the markets of network cameras and video encoders. They develop embedded software for security cameras. The cameras are designed to be accessible through APIs that are developed and maintained by the company.

@@

\section{Methodology} \label{methodology}
This study has been conducted using the case study meth-odology \cite{runeson2009guidelines}. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API interface and API documentation have been used as units of analysis. The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case, robson2002real, runeson2009guidelines, yin2013case}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of drivers of API change and that input from the industry is vital to the success of the research.

@@

The study was conducted in two phases. The aim of the first phase was to generate hypotheses and to formulate clear research questions. This was achieved by using a data-driven approach to analysing the interface code and documentation. The second phase included analysis of the interviews, during which we aimed to confirm the hypotheses that were formulated in the first phase. Both phases were performed iteratively. This created the opportunity to improve the data analysis as the study progressed, as well as, allowing us to adapt to possible changes in direction, due to the hypothesis generating approach of the first phase. An overview of the methodology can be found in Table \ref{table:meth_steps}.

@@

%How the initial inspection was performed
The API interface code was inspected sequentially, starting with v1.4. Both versions of the API underwent the same type of inspection. This involved extracting method signatures, enums, structs, typedefs and macros, which were subsequently stored in spreadsheets. The extraction involved a lot of manual work, but Git's \cite{git} diff command was used to identify changes in files which exist in both versions. After a change was identified in the code, a description of the corresponding change in the documentation was added to the spreadsheet. The module and file in which the change was identified was then logged together with an ID and a description of the change. The data collection from code and documentation was performed independently by both researchers to reduce the risk of human errors affecting the results. If the data collected by each researcher differed, the cause of the discrepancy was investigated and resolved.

@@

%Codes
To analyse the data, each change identified in the spreadsheet was coded based on which module the change occurred in and based on if the change added, removed, modified or did not affect functionality. 
%Concepts
Grouping the codes by module allowed us to form general concepts of change. However, changes that affected multiple modules were classified on their own.
%Categories
These concepts were then grouped into categories which were based on common patterns between the concepts. An example of a pattern found during the analysis was that several concepts followed the pattern of adding new functionality.    
%Theory
These categories were then used to identify the most notable trends in the evolution of the API. These trends were later used as a basis for discussion during the interview.

%RQ1
The trends identified previously were then classified with the help of the decision tree used for classifying the types of software evolution and software maintenance \cite{chapin2001types}. Each of the previously identified trends were given a main type of change and, if appropriate, a secondary type. Main types of change identify the primary motive for a trend, whereas the secondary types are additional motivations for a trend. For example, if an identified trend was the re-implementation of an old module that included new functionality and made use of a new interface, it would be classified as mainly enhancive and secondarily groomative. In addition to this, we analysed the trends based on Lehman's Laws, to determine if they applied to the changes made to the API. This information, as well as the results of the interview, were then used as the basis for answering \textbf{RQ1}.

@@

\smallskip \noindent
\textbf{Updative  }
%Define
Updative changes are defined to be changes to documentation of a system that are made to conform with changes in the code. 
%Data and example
A single change was determined to be updative; updating the API documentation to include the new functionality of v2.0. The updates included a new API specification, deprecation list and the inclusion of the new features and libraries.
%Conclusion
Updative changes have very little impact on the functionality of the API yet they are still important changes for the usability of the API \cite{shi2011empirical}. Changes to the API cause the need for updative changes to its documentation, without them, the usability of the API could suffer. We therefore conclude that the evolution of this API was not driven by updative changes.

@@

\smallskip \noindent
\textbf{Reformative  }
%Define
Reformative changes are defined to be \newline changes that update the documentation to the stakeholders needs. An example of such a change is that the structure of the documentation is changed. 
%Data and Example
One reformative change was identified between v1.4 and v2.0; a redesign of the existing example programs. The aim of this change was to update the examples to illustrate the new functionality in the API, showing how the modules interact and at the same time, presenting a standard for how to use the API.
%Conclusion
We therefore conclude that reformative changes, while important for the usability of the API, did not drive the evolution of the API.

@@

\smallskip \noindent
\textbf{Conservation of Familiarity  } 
The documentation that is included with the API was rigorously updated with the introduction of v2.0. All functionality which was added or modified in v2.0 was updated accordingly in the documentation. The same pattern can be seen for the example code, which went through major changes as a result of the changes to the API interface. Company A's intent to control the way that their API is used also suggests that the knowledge of how to use the API is of importance, since if the API is used in too many different ways, it will be more difficult for company A to have relevant documentation and example code. It was also expressed during the interview that further transparency in how the API should be used could be achieved. One example mentioned of how this could be achieved is by improving the expressiveness regarding which API functions are supported on which hardware types. These facts support that the law of Conservation of Familiarity played an important part in the evolution of the API.

@@

\smallskip \noindent
\textbf{Declining Quality  } 
The interviews established that a groo-mative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. Our analysis also shows that changes related to quality are bound to happen, as other types of changes often incorporate quality related aspects. We therefore settle that the law of Declining Quality was a driving force for the evolution of the API.

@@

\subsection{Reverse Engineering Change Decisions}
After cross-examining our initial analysis with the interviewee's responses, we found that 85\% of our main type predictions matched the actual motive of the change and that 85\% of our secondary type predictions were correct. The kappa value \cite{cohen1968weighted} for the predictions was 0.812, which shows that there is a high level of agreement between the predictions and the change decisions. In Table \ref{table:reverse_engineering}, a detailed view of each prediction juxtaposed against the actual change decision can be found for both the main and secondary types.

@@

\subsection{Implications for Research} 
The implications of our findings should encourage research to further study APIs by relating them to existing software evolution theories. Considering that we were able to successfully reverse engineer the changes made to the API in this study, it opens up possibilities to base future research on, for example, the types of software evolution and software maintenance. Since the kappa analysis showed that there was a high level of agreement between our classification and the actual motives to the change, it shows that the types of software evolution and software maintenance could even be used as the basis for a large-scale quantitative analysis to map the types of changes made to APIs.

@@

One way of finding the solution to this problem would be by in a more structured way including the application developers in the feedback loop. Since the application developers are on the receiving end of the changes, we believe that their needs have to be assessed before the problem can be solved. With our study we identified both that the law of Feedback System is of great importance to company A and that it is one of the key areas of API development that company A wishes to improve. One of the challenges posed by software evolution is that the activities carried out are often intangible and non-objective \cite{chapin2001types, lientz1980software}. Based on this, we propose that our study which has granulated the activities is used as a basis for communicating the current stage of the evolution of their API. We hope that this classification can be used in its current form, to help quantify the activities and to provide an overview of them. We also hope that the approach we used can be used to predict which types of activities should be emphasised in the future evolution of the API. With our study we have also demonstrated that the classification of changes is achievable, which, in turn, should encourage company A to adopt this method of classifying change.

\section{Future Work} \label{conclusion}
With this study, we conclude that studying API evolution with the help of existing theories, provides a new and structured approach to the field of API evolution. Our study shows that functionality is the largest driving force of change in the evolution of the studied API. It also suggests that changes to APIs occur infrequently and sporadically. This indicates that the law of Conservation of Organisational Stability did not play a significant role in the evolution of this API. However, given that this study was only performed on a single API, we cannot generalise this finding. Furthermore, we have identified the importance of having a structured feedback loop when maintaining an API. We have also provided company A with a foundation for facilitating such a change.
